---
title: "Notes on Tidyverse"
author: "Shashank Sule"
date: "9/15/2019"
output: pdf_document
header-includes:
  -\usepackage[margin=1in]{geometry}
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Tidyverse is the collection of tools which help you visualize and explore data. The example we will use is the gapminder data set. 

```{r}
require(gapminder)
require(dplyr)

gapminder
```

Note that `gapminder` is a `tibble`, a bit different from `data.frame`. 

# Verbs from `dplyr`

Verbs are the functions within `dplyr`. Every time you use a verb, you use a pipe which composes verbs. 

1. `filter`

`filter` literally filters your data frame according to the criteria you give. Suppose you want to filter the dataset for 2007. 

```{r}
gapminder %>% 
  filter(year == 2007) #Indexing the array by boolean variables

#You can also double filter

gapminder %>% filter(year == 2007, country == "United States")

```

2. `arrange`

`arrange` sorts the dataset according to the column you want to arrange by. By default it arranges according to ascending order. To sort in descending order you do `arrange(desc(#colname))`

```{r}
# Sort in ascending order of lifeExp
gapminder %>% arrange(lifeExp)
  
# Sort in descending order of lifeExp
gapminder %>% arrange(desc(lifeExp))
```


3. `mutate`

`mutate` either changes an existing column in the data frame or adds a new one. Note that it is not a "static" function: it does not make any changes to the data frame you pass as an argument. Instead, it returns a new data frame that is a result of passing the original data frame through `mutate`. 

```{r}
# Use mutate to change lifeExp to be in months
gapminder %>% mutate(lifeExp = lifeExp * 12)

# Use mutate to create a new column called lifeExpMonths

gapminder %>% mutate(lifeExpMonths = lifeExp * 12)
```

```{r}
# Filter, mutate, and arrange the gapminder dataset

gapminder %>%
  filter(year == 2007) %>%
  mutate(lifeExpMonths = 12 * lifeExp)%>%
  arrange(desc(lifeExpMonths))
```

4. `summarize`

`summarize` returns a `tibble` with "summarizing" statistics of the dataset you give. 

```{r}
# Summarize to find the median life expectancy

gapminder %>% summarize(medianLifeExp = median(lifeExp))


# Filter for 1957 then summarize the median life expectancy
gapminder %>% filter(year == 2007) %>%
 summarize(medianLifeExp = median(lifeExp))

# Summarize can even return two values

gapminder %>% filter(year == 1957) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))

```

5. `group_by`

In the previous example, filter allowed you to select a value of one particular column, which then you summarized. In order to summarize over all values of a particular column, we use `group_by`. It's equivalent to using a for loop on `filter` where the loop runs over the all the possible arguments passable to filter (which is what you would do to get an overall summary table if you didn't know about `group_by`).

```{r}
# Find median life expectancy and maximum GDP per capita in each year
gapminder %>%
  group_by(year) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))


# Find median life expectancy and maximum GDP per capita in each continent in 1957

gapminder %>%
 filter(year==1957) %>%
 group_by(continent) %>%
 summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))

#You can even pass two arguments to group_by. In that case it returns the summary table over all possible combinations of the values of the passed arguments 

gapminder %>%
 group_by(continent, year) %>%
 summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))

# In the above output, summarize gave the summaries at all possible (year, continent) combinations. 
```


# Visualization with ggplot2

`ggplot2` is the canonical data visualization tool. We'll load it here

```{r}
library("ggplot2")
```

The basic syntax for making a plot is `ggplot(#dataframe, ae(x=#xcomponent, y=#component))`. The `+` operator will signify additional features to the plot. 

1. Scatter plot

Scatter plots are achieved using `+ geom_point()`. 
```{r, fig.height=3, fig.width=4, fig.align='center'}
library(dplyr)
library(gapminder)
gapminder_1952 <- gapminder %>%
  filter(year == 1952)
ggplot(gapminder_1952, aes(x = pop, y = gdpPercap)) +
  geom_point()
ggplot(gapminder_1952, aes(x=pop, y=lifeExp)) + geom_point() 

```

2. Log scale

You can use log axes by adding `+ scale_x_log10()` or `+ scale_y_log10()`. 

```{r, fig.height=3, fig.width=4, fig.align='center'}
ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) +
  geom_point() + 
  scale_x_log10()
ggplot(gapminder_1952, aes(x=pop, y=gdpPercap)) + 
  geom_point() + 
  scale_x_log10() +
  scale_y_log10()

```

3. Aesthetics

You can add more than just `x` and `y` in the `aes()` parameter in `ggplot`. For example, you can add `color` which is the colour of the data points, and `size` which is their size. For colour, ggplot automatically adds a legend! 

```{r, fig.height=3, fig.width=4, fig.align='center'}
ggplot(gapminder_1952, aes(x=pop, y=lifeExp,color=continent)) + 
  geom_point() + 
  scale_x_log10()

ggplot(gapminder_1952, aes(x = pop, y = lifeExp, color = continent, size = gdpPercap)) +
  geom_point() +
  scale_x_log10()
```


Question: do `color` and `size` only matter when you use the scatter plot? Question: is it possible to modify the legend (say represent the `gdpPercap` values in log scale?)

4. Faceting

Faceting refers to the practice of making multiple plots within the same figure. In `ggplot` you can make multiple plots according to their indices in the dataset. To facet a plot, we use `+ facet_wrap(~#indexingparameter)`. 

```{r, fig.height=3, fig.width=4, fig.align='center'}
# Here we break down the analysis of population vs life expectancy according to continent. Conveniently, continent is a column in our data frame. 

# Scatter plot comparing pop and lifeExp, faceted by continent
ggplot(gapminder_1952, aes(x=pop, y=lifeExp, size = pop, color = continent)) + 
geom_point() + 
scale_x_log10() + 
facet_wrap(~ continent)
```


Note that we can apply the `color` and the `facet_wrap` functions to the same parameters. The result is obviously that all the colours go into the same subplot. 

```{r, fig.height=3, fig.width=4, fig.align='center'}
# Scatter plot comparing gdpPercap and lifeExp, with color representing continent
# and size representing population, faceted by year
ggplot(gapminder, aes(x=gdpPercap, y=lifeExp, color = continent, size = pop)) + 
geom_point() + 
scale_x_log10() + 
facet_wrap(~ year)
```

5. Axes limits 

Set axes limits using `+ expand_limits(y=#value)`

```{r, fig.height=3, fig.width=4, fig.align='center'}

# Here we'll visualize some summarized data
by_year <- gapminder %>%
  group_by(year) %>%
  summarize(medianLifeExp = median(lifeExp),
            maxGdpPercap = max(gdpPercap))

# Create a scatter plot showing the change in medianLifeExp over time

ggplot(by_year, aes(x = year, y = medianLifeExp)) + 
  geom_point() + 
  expand_limits(y=0)

# Summarize medianGdpPercap within each continent within each year: by_year_continent
by_year_continent <- gapminder %>%
 group_by(year, continent) %>%
 summarize(medianGdpPercap = median(gdpPercap))

# Plot the change in medianGdpPercap in each continent over time

ggplot(by_year_continent, aes(x=year, y=medianGdpPercap, color = continent)) + 
geom_point() + 
expand_limits(y=0)
```

6. Line plot

To make a Line plot, use `+ geom_line()`

```{r, fig.height=3, fig.width=4, fig.align='center'}
# Summarize the median gdpPercap by year, then save it as by_year
by_year <- gapminder %>%
  group_by(year) %>%
  summarize(medianGdpPercap = median(gdpPercap))

# Create a line plot showing the change in medianGdpPercap over time

ggplot(by_year, aes(x=year, y=medianGdpPercap)) + 
geom_line() + 
expand_limits(y=0)

# Summarize the median gdpPercap by year & continent, save as by_year_continent
by_year_continent <- gapminder %>%
  group_by(year, continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))

# Create a line plot showing the change in medianGdpPercap by continent over time

ggplot(by_year_continent, aes(x=year, y=medianGdpPercap, color = continent)) + 
geom_line() + 
expand_limits(y=0)
```

7. Bar plot

Use `+ geom_col()`

```{r, fig.height=3, fig.width=4, fig.align='center'}
# Summarize the median gdpPercap by year and continent in 1952
by_continent <- gapminder %>%
  filter(year==1952) %>%
  group_by(continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))

# Create a bar plot showing medianGdp by continent

ggplot(by_continent,aes(x=continent, y=medianGdpPercap)) + 
geom_col()
```


8. Histogram 

Use `+ geom_histogram(binwidth=#setwidth)`. Note that for hist, you need only specify `x` in `aes()`. 

```{r, fig.height=3, fig.width=4, fig.align='center'}
gapminder_1952 <- gapminder %>%
  filter(year == 1952)

# Create a histogram of population (pop), with x on a log scale

ggplot(gapminder_1952, aes(x=pop)) + 
scale_x_log10()+
geom_histogram()
```


9. Box-Whisker plot

Use `+ geom_boxplot()` 

```{r, fig.height=3, fig.width=4, fig.align='center'}
gapminder_1952 <- gapminder %>%
  filter(year == 1952)

# Create a boxplot comparing gdpPercap among continents

ggplot(gapminder_1952, aes(x=continent, y=gdpPercap)) + 
geom_boxplot() + 
scale_y_log10()
```

10. Adding title and x and y labels 

For title use `+ ggtitle()` and for labels use `labs(x=#namehere, y=#namehere)`

```{r, fig.height=3, fig.width=4, fig.align='center'}
gapminder_1952 <- gapminder %>%
  filter(year == 1952)

# Add a title to this graph: "Comparing GDP per capita across continents"
ggplot(gapminder_1952, aes(x = continent, y = gdpPercap)) +
  geom_boxplot() +
  scale_y_log10() + 
  ggtitle("Comparing GDP per capita across continents") + 
  labs(x="Continent", y="GDP per Capita")
```

