---
title: "Case Study 1"
author: "Sarah Boese and Shashank Sule"
header-includes:
   - \usepackage{amsmath}
   - \usepackage{amsthm}
   - \usepackage{amssymb}
   - \usepackage{amsfonts}
   - \usepackage[margin=1in]{geometry}
output: pdf_document
fontsize: 12pt
date: "07/11/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exploring the dataset

```{r, echo=FALSE}
install.packages("ggplot2")
install.packages("dplyr")
require(ggplot2)
require(dplyr)
library(readr)
ItalyMarriageRates <- read_csv("ItalyMarriageRates.csv")
View(ItalyMarriageRates)
Dats <- ItalyMarriageRates %>% mutate(Period = if_else(Year < 1940, 1, if_else(Year > 1944, 3, 2)))
ggplot(Dats, aes(x=Dats$Year,y=Dats$`Rates (per 1000)`, color=Dats$Period))+
  geom_point()
```

It seems that marriage rates were lower per 1000 during the world war. We model them using a Hierarchical Poisson model (akin to what was seen in the midterm exam). Note that we regard the inter-war years as 1940-1944 instead of 1939-1945 because the second world war began and culminated in the middle of the two years 1939 and 1945 respectively (so there could have been a mixture of pre-war, inter-war, and post-war rates in '39 and '45). 

# Model

Let $Y_{i}$ be the marriage rate in year $i$ and $Y_{i} \sim \text{Poisson}(\lambda_i)$ where 

$$ \lambda_i = \begin{cases}  \lambda_1 & 1936 \leq i \leq 1938 \\
                              \lambda_2 & 1939 \leq i \leq 1945 \\
                              \lambda_3 & 1946 \leq i \leq 1951 \end{cases}$$
                              
$$ \lambda_j \sim \text{Gamma}(a_j, b_j) \,, j:1 \leq j \leq 3$$

We'll introduce a common normal hyperprior for $a_j$, $b_j$ each where the means reflect the shape and rate parameters deduced from the mean and variance from the data. 

```{r}
col <- Dats %>% select(`Rates (per 1000)`)
mean(col$`Rates (per 1000)`)
(sd(col$`Rates (per 1000)`))^2
```

So $a/b = 7.3125$ and $(a/b)*(1/b) = 7.3125*(1/b) = 1.9625$ so the value for $b$ is 

```{r}
7.3125/1.9625
```

and consequently $a$ is 

```{r}
3.726115*7.3125
```

Thus, $a_j \sim \text{N}(27.24722, 1)$ and $b_j \sim \text{N}(3.726115,1)$. 

# Simulation 

Data: 

```{r}
N <- length(ItalyMarriageRates$Year)
Periods <- Dats$Period
Y <- ItalyMarriageRates$`Rates (per 1000)`
```

Model: 

```{r}

modelString <-"
model {
## likelihood
for (i in 1:N){
y[i] ~ dpois(lambda[Periods[i]])
}

## priors
for (j in 1:3){
lambda[j] ~ dgamma(a,b)
}

## hyperpriors
a ~ dnorm(A,100)
b ~ dnorm(B,100)
}
"
```

Setting up JAGS: 

```{r}
the_data <- list("y" = Y, "N" = 16, "A" = 27.24722, "B" = 3.726115, "Periods" = Periods)

initsfunction <- function(chain){
  .RNG.seed <- c(1,2)[chain]
  .RNG.name <- c("base::Super-Duper",
                 "base::Wichmann-Hill")[chain]
  return(list(.RNG.seed=.RNG.seed,
              .RNG.name=.RNG.name))
}
```

Running the simulation: 

```{r}
posterior <- run.jags(modelString,
                      n.chains = 5,
                      data = the_data,
                      monitor = c("lambda", "a", "b"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000,
                      thin = 5)
```

Summary and Diagnostics: 

```{r}
summary(posterior)
```

```{r}
plot(posterior, vars = "lambda[2]")
plot(posterior, vars = "lambda[1]")
plot(posterior, vars = "lambda[3]")
```

We ran 5 chains with a thinning value of 5 to ensure exploration of parameter space alongside low correlation. 

# Commentary

From the summary of posterior data in the poisson model, the mean inter-war rate was 1 point lower than the rates found post-war and pre-war. However, notice that the standard deviation for the post-war rate is almost 1, which means that when we actually sample from the posterior, 16% of the time the inter-war rate will be in the parameter space of the pre-war rate. Thus, we cannot satisfactorily claim that the inter-war rate was different from the post war one. 
