---
title: "Case Study 3"
authors: "Olivia Fiol and Shashank Sule"
date: "04/12/2019"
header-includes:
   - \usepackage{amsmath}
   - \usepackage{amsthm}
   - \usepackage{amssymb}
   - \usepackage{amsfonts}
   - \usepackage[margin=1in]{geometry}
output: pdf_document
fontsize: 12pt
---

Our latent class model has the likelihood function
\begin{align*}
  Y_i=y_i|z_i,p_{z_i}\sim\text{Binomial}(m,p_{z_i}),
\end{align*}
with 
\begin{align*}
z_i|\pi\sim\text{Bernoulli}(\pi).
\end{align*} The prior distributions are the following:
\begin{align*}
&p_1|z_i=1\sim\text{Uniform}(0.4,0.6),\\
&p_0|z_i=0\sim\text{Uniform}(0.5,1), \text{ and}\\
&\pi\sim\text{Beta}(a,b).
\end{align*}


# Simulation 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(gridExtra)
require(ProbBayes)
require(tidyverse)
require(runjags)
require(coda)
require(devtools)
crcblue <- "#2905a1"
knitr::opts_chunk$set(echo = TRUE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

```{r}
library(readr)
TrueFalseScores <- read_csv("TrueFalseScores.csv")
View(TrueFalseScores)
```



```{r}
modelString<-"
model {
## sampling
for (i in 1:N){
   theta[i] <- equals(z[i], 1) * p1 + equals(z[i], 0) * p0
y[i] ~ dbin(theta[i], m)
}
for (i in 1:N){
   z[i] ~ dbern(pi)
}
## priors
p1 ~ dbeta(1,1)T(0.4,0.6)
p0 ~ dbeta(1,1)T(0.5, 1)
pi ~ dbeta(a,b)
}
"
```

```{r}
y <- TrueFalseScores$Score
N <- length(y)

initsfunction <- function(chain){
  .RNG.seed <- c(1,2)[chain]
  .RNG.name <- c("base::Super-Duper",
                 "base::Wichmann-Hill")[chain]
  return(list(.RNG.seed=.RNG.seed,
              .RNG.name=.RNG.name))
}

the_data <- list("y" = y, "N" = N, "m"=40, "a"=2, "b"=4)
```

```{r, message=FALSE, warning=FALSE, results="hide"}
posterior <- run.jags(modelString,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("p1", "p0", "theta"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000,
                      thin = 10, 
                      inits = initsfunction)
```

```{r}
summary(posterior)
```

```{r, echo=FALSE}
vars_list <- c("theta[1]", "theta[10]", "p0", "p1")
plot(posterior, vars=vars_list)
```

The chain exhibits MCMC convergence; note that the `theta` values for the first five observations are approximately `0.43` and the values for the next 10 are approximately `0.83`. The data thus leads us to the conclusion that there was a guessing group. 